DCM_investigation

### 2020-11-12 article_figures_ARTHUR_CAPET.Rmd
Light Plots + chla/bbp ratio

p2b_start = c(PAR0 = 1500 ,  ks0 = 0.04 , kl0=0.001) 
p2b_lower = c(PAR0 = 0    ,  ks0 = 0    , kl0=0    )
p2b_upper = c(PAR0 = 2500 ,  ks0 = 3   , kl0=10)

@@

## First attempt derives APR @ 1m from interpolation and look for 1/100 of this value
isoPAR <- function(depth, PAR){
  require(pracma)
  # Intended for Casted DF
  if (FilterForPAR(depth,PAR)){
    out    <- try_default({
    d <- depth[which( !is.na(depth) & PAR>0.1  )]
    P <- PAR[which( !is.na(depth) & PAR>0.1  )]
             p1 <-interp1(d,P, 1)      
      z100 <- interp1(P,d, p1/100) 
      return(z100)},
      DefOut_isoPAR)
  }else{
    out <- DefOut_isoPAR
  }
  return(out)
}

## Second option 
isoPARb <- function(depth, PAR){
  require(pracma)
  # Intended for Casted DF
  if (FilterForPAR(depth,PAR)){
    out    <- try_default({
    d <- depth[which( !is.na(depth) & PAR>0.1  )]
    P <- PAR[which( !is.na(depth) & PAR>0.1  )]

    m <<-nls(PAR ~ Att_2band(depth,PAR0,ks0,kl0),
           data = data.frame(depth=depth, PAR=PAR),
           start = p2b_start, 
           lower = p2b_lower,
           upper = p2b_upper,
           algorithm = "port", 
           control = list(maxiter= 1000)
           )
    mcoef  <<- coef(m)
    PAR0   <-mcoef[1]
    PARcon <- predict(m,data.frame(depth=seq(1,200)))    
    z100 <- interp1(PARcon,seq(1,200), PAR0/100) 
      return(z100)},
      DefOut_isoPAR)
  }else{
    out <- DefOut_isoPAR
  }
  return(out)
}

@@

if (freshstart){
integrated_diags_par <- ddply(initPAR, .(juld,platform), summarize,
                          Ziso1 = isoPAR(depth = depth , PAR = par),
                          Ziso1b = isoPARb(depth = depth , PAR = par)
                          )
}

@@

```{r}
# Add scalr diagnostics to the complete dataframe for vertical integration
if (freshstart){
complete_tmp2 <- merge(complete_tmp,integrated_diags_par, by=c('juld','platform'), all.x = TRUE)
}

```{r add light data to full data}
interpolated_light_data <- ldply(as.list(unique(complete_tmp2$juld)), function(i){
  tmp1 <- complete_tmp2[complete_tmp2$juld == i,]
  tmp2 <- initPAR[initPAR$juld == i,]
  if(nrow(tmp2) == 0){
    absolute_light <- rep(NA, nrow(tmp1))
    # relative % of incoming surface PAR
    relative_light <- absolute_light
  }else{
    absolute_light <- approx(tmp2$depth, tmp2$par, tmp1$depth)$y
    # compute relative % of incoming surface PAR
    relative_light <- (absolute_light/tmp2$par[1])*100 # could also put max(tmp2$par) car surface = max_par, therefore no ordering issue will appear (but tmp2$par[1] is good) 
  }
  data.frame(depth = tmp1$depth, absolute_light = absolute_light, relative_light = relative_light)
})

# add it to complete_tmp2
complete_tmp2 <- cbind(complete_tmp2, light = interpolated_light_data$relative_light, light_absolute = interpolated_light_data$absolute_light)

@@

                            #Ziso1         = isoPAR(depth = depth , PAR = par),

@@

                            ISOlow       = try_default(approx(x=depth, y=light, xout=ZLow, rule=2)$y,NA, quiet = FALSE),
                            ISO90B       = try_default(approx(x=depth, y=light, xout=Z90B, rule=2)$y,NA, quiet = FALSE),
                            ISO90T       = try_default(approx(x=depth, y=light, xout=Z90T, rule=2)$y,NA, quiet = FALSE),
                            ISOMLD       = min(light, na.rm = TRUE),
                            ISOMAX       = try_default(approx(x=depth, y=light, xout=unique(Zmax), rule=2)$y,NA, quiet = FALSE),
                            ISODEMI      = try_default(approx(x=depth, y=light, xout=unique(Zdemi), rule=2)$y,NA, quiet = FALSE),
                          # just to see
                            ISO_iso1 = try_default(approx(x=depth, y=light, xout=unique(Ziso1b), rule=2)$y,NA, quiet = FALSE)

@@

dbdf$fluo_norm <- dbdf$fluo/dbdf$chla_dcm
dbdf$sigma_norm <- (dbdf$sigma-dbdf$sigmaMLD)/(dbdf$sigmaMAX-dbdf$sigmaMLD+.00000001)
dbdf$depth_norm <- (dbdf$depth-dbdf$MLD)/(dbdf$Zmax-dbdf$MLD+.00000001)
#dbdf$depth_norm_layers<-factor(dbdf$depth_norm_layers,levels=rev(levels(dbdf$depth_norm_layers)))

@@

dbdf$sigma_norm_layers <- cut(dbdf$sigma_norm,seq(0.05,3.05,.05))
dbdf$depth_norm_layers <- cut(dbdf$depth_norm,seq(0.05,3.05,.05))
```

```{r}
blbl<-subset(dbdf, select=c("sigma_norm_layers","fluo_norm", "depth_norm_layers","depth_norm","sigma_norm","bbp_norm1","month",'lon'))
require(reshape2)
blbl<-melt(blbl, id.vars=.(sigma_norm_layers,depth_norm_layers,month,depth_norm,sigma_norm,lon))

@@

blbl$variable<-as.character(blbl$variable)
blbl$variable[which(blbl$variable=="fluo_norm")]<-"Chla - norm."
blbl$variable[which(blbl$variable=="bbp_norm1")]<-"Bbp - norm."

blbl$variable<-factor(blbl$variable, levels=c("Chla - norm.","Bbp - norm."))

bigplot1<-ggplot(subset(blbl, month %in% c(3,4,5,6)), aes(x=depth_norm, y=value))+
      geom_boxplot(aes(group=depth_norm_layers), outlier.size=0.5)+
  theme_bw()+coord_flip(xlim = c(0,2), ylim=c(0,2))+
  facet_grid(monthn~variable)+
  geom_vline(xintercept  = 1)+xlab('Depth - norm.')+ylab('')

bigplot2<-ggplot(subset(blbl, month %in% c(7,8,9,10)), aes(x=depth_norm, y=value))+
      geom_boxplot(aes(group=depth_norm_layers), outlier.size=0.5)+
  theme_bw()+coord_flip(xlim = c(0,2), ylim=c(0,2))+
  facet_grid(monthn~variable)+
  geom_vline(xintercept  = 1)+xlab('Depth - norm.')+ylab('')
```

@@

ggplot(tmp_sig, aes(x=month, fill=variable, color=variable)) +
    geom_smooth(aes(y=value),stat = 'summary',
#              fun.data = median_hilow,
              fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), alpha=0.5)+
  geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2, notch = TRUE)+
        scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab('[kg/m3]')+theme_bw()+coord_cartesian(ylim=c(16,10))+scale_y_reverse()+
      scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal,
                       name="Density Horizons", 
                       labels=list(TeX('$\\sigma_{MLD}$'),
                                TeX('$\\sigma_{DCM}$'),
                                TeX('$\\sigma_{low}$'),
                                TeX('$\\sigma_{50,bottom}$'),
                                TeX('$\\sigma_{50,top}$'),
                                TeX('$\\sigma_{\\mathrm{PAR}\\,1%}$')))+
  geom_hline( yintercept = 14.5)

@@

tmp_iso1 <- subset(tmp_fitco, select = c('month','ISOMLD','ISOMAX','ISOlow','ISO90B','ISO90T','ISO_iso1' ,'dcm_or_not','year','shape'))

tmp_iso1$sigmaMAX[tmp_iso1$dcm_or_not==0]<-NA
tmp_iso1$sigmaMAX[tmp_iso1$Zmax>200]<-NA

tmp_iso <- melt(tmp_iso1[,-which(colnames(tmp_iso1)=='dcm_or_not')], id.vars=c('month','year','shape'))

ggplot(tmp_iso, aes(x=month, fill=variable, color=variable)) +
    geom_smooth(aes(y=value),stat = 'summary',
#              fun.data = median_hilow,
              fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), alpha=0.5)+
  geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2, notch = TRUE)+
        scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  #scale_y_continuous(trans = 'log10') + ylab('[microMoleQuanta/m^2/sec]') +
  ylab('% of Incoming PAR') + theme_bw()  + scale_y_continuous(trans = 'log10') +
      scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal,
                       name="ISO Horizons", 
                       labels=list(TeX('$\\ISO_{MLD}$'),
                                TeX('$\\ISO_{DCM}$'),
                                TeX('$\\ISO_{low}$'),
                                TeX('$\\ISO_{50,bottom}$'),
                                TeX('$\\ISO_{50,top}$'),
                                TeX('$\\ISO_{\\mathrm{PAR}\\,1%}$')))

>>>>

p2b_start = c(PAR0 = 1500 ,  ks0 = 0.04 , kl0=0.01) 
p2b_lower = c(PAR0 = 0    ,  ks0 = 0    , kl0=0    )
p2b_upper = c(PAR0 = 2500 ,  ks0 = 3   , kl0=10)

@@

## First attempt derives PAR @ 1m from interpolation and look for 1/100 of this value
isoPAR <- function(depth, PAR){
  require(pracma)
  # Intended for Casted DF
  if (FilterForPAR(depth,PAR)){
    out    <- try_default({
    d <- depth[which( !is.na(depth) & PAR>0.1  )]
    P <- PAR[which( !is.na(depth) & PAR>0.1  )]
             p1 <-interp1(d,P, 1)      
      z100 <- interp1(P,d, p1/100) 
      return(z100)},
      DefOut_isoPAR)
  }else{
    out <- DefOut_isoPAR
  }
  return(out)
}

## Second option 
isoPARb <- function(depth, PAR){
  require(pracma)
  # Intended for Casted DF
  if (FilterForPAR(depth,PAR)){
    out    <- try_default({
    d <- depth[which( !is.na(depth) & PAR>0.1  )]
    P <- PAR[which( !is.na(depth) & PAR>0.1  )]
    m <<-nls2(PAR ~ Att_2band(depth,PAR0,ks0,kl0),
           data = data.frame(depth=depth, PAR=PAR),
           start = p2b_start, 
           lower = p2b_lower,
           upper = p2b_upper,
           algorithm = "port", 
           control = list(maxiter= 1000)
           )
    mcoef  <<- coef(m)
    PAR0   <-mcoef[1]
    PARcon <- predict(m,data.frame(depth=seq(1,200)))    
    z100 <- interp1(PARcon,seq(1,200), PAR0/100) 
      return(c(z100,PAR0))},
      DefOut_isoPAR)
  }else{
    out <- DefOut_isoPAR
  }
  return(out)
}

@@

if (freshstart){
integrated_diags_par <- ddply(initPAR, .(juld,platform), summarize,
                          Ziso1 = isoPAR(depth = depth , PAR = par),
                          Ziso1b = isoPARb(depth = depth , PAR = par)[1],
                          PAR0  = isoPARb(depth = depth , PAR = par)[2]
                          )
}

@@

```{r}
# Add scalr diagnostics to the complete dataframe for vertical integration
if (freshstart){
complete_tmp2 <- merge(complete_tmp,integrated_diags_par, by=c('juld','platform'), all.x = TRUE)
}
```

```{r add light data to full data}
complete_tmp2 <- ddply(complete_tmp2,c('juld','platform'), function(dsub){
  tmp <- subset(initPAR, (platform==dsub$platform[1]) & (juld==dsub$juld[1]) )
  if(nrow(tmp) == 0){
    absolute_light <- rep(NA, nrow(dsub))
    relative_light <- absolute_light
  }else{
    absolute_light <- approx(tmp$depth, tmp$par, dsub$depth)$y
    relative_light <- (absolute_light/dsub$PAR0)*100 # could also put max(tmp2$par) car surface = max_par, therefore no ordering issue will appear (but tmp2$par[1] is good) 
  }
  dsub$absolute_light<-absolute_light
  dsub$relative_light<-relative_light
  return(dsub)
})

@@

@@

                            ISOlow       = try_default(approx(x=depth, y=relative_light, xout=ZLow, rule=2)$y,NA, quiet = FALSE),
                            ISO90B       = try_default(approx(x=depth, y=relative_light, xout=Z90B, rule=2)$y,NA, quiet = FALSE),
                            ISO90T       = try_default(approx(x=depth, y=relative_light, xout=Z90T, rule=2)$y,NA, quiet = FALSE),
                            ISOMLD       = try_default(approx(x=depth, y=relative_light, xout=unique(MLD), rule=2)$y,NA, quiet = FALSE),
                            ISOMAX       = try_default(approx(x=depth, y=relative_light, xout=unique(Zmax), rule=2)$y,NA, quiet = FALSE),
                            ISODEMI      = try_default(approx(x=depth, y=relative_light, xout=unique(Zdemi), rule=2)$y,NA, quiet = FALSE),
                            ISO_iso1     = try_default(approx(x=depth, y=relative_light, xout=unique(Ziso1b), rule=2)$y,NA, quiet = FALSE),
                          # Add ABSOLUTE Light levels
                            ISOalow       = try_default(approx(x=depth, y=light_absolute, xout=ZLow, rule=2)$y,NA, quiet = FALSE),
                            ISOa90B       = try_default(approx(x=depth, y=light_absolute, xout=Z90B, rule=2)$y,NA, quiet = FALSE),
                            ISOa90T       = try_default(approx(x=depth, y=light_absolute, xout=Z90T, rule=2)$y,NA, quiet = FALSE),
                            ISOaMLD       = try_default(approx(x=depth, y=light_absolute, xout=unique(MLD), rule=2)$y,NA, quiet = FALSE),
                            ISOaMAX       = try_default(approx(x=depth, y=light_absolute, xout=unique(Zmax), rule=2)$y,NA, quiet = FALSE),
                            ISOaDEMI      = try_default(approx(x=depth, y=light_absolute, xout=unique(Zdemi), rule=2)$y,NA, quiet = FALSE),
                            ISOa_iso1     = try_default(approx(x=depth, y=light_absolute, xout=unique(Ziso1b), rule=2)$y,NA, quiet = FALSE)

@@

dbdf$fluo_norm <- dbdf$fluo/dbdf$chla_dcm

dbdf$sigma_norm <- (dbdf$sigma-dbdf$sigmaMLD)/(dbdf$sigmaMAX-dbdf$sigmaMLD+.00000001)
dbdf$depth_norm <- (dbdf$depth-dbdf$MLD)/(dbdf$Zmax-dbdf$MLD+.00000001)
#dbdf$depth_norm_layers<-factor(dbdf$depth_norm_layers,levels=rev(levels(dbdf$depth_norm_layers)))

@@

dbdf$ration <- dbdf$fluo_norm/ dbdf$bbp_norm1 
#dbdf$ratio  <- dbdf$fluo / dbdf$bbp

dbdf$sigma_norm_layers <- cut(dbdf$sigma_norm,seq(0.05,3.05,.1))
dbdf$depth_norm_layers <- cut(dbdf$depth_norm,seq(0.05,3.05,.1))
```

```{r}
blbl<-subset(dbdf, select=c("sigma_norm_layers","fluo_norm", "depth_norm_layers",
                            "depth_norm","sigma_norm","bbp_norm1","month",'lon',"ration"))
require(reshape2)

blbl<-melt(blbl, id.vars=.(sigma_norm_layers,depth_norm_layers,month,depth_norm,sigma_norm,lon))

@@

blbl$variable<-as.character(blbl$variable)
blbl$variable[which(blbl$variable=="fluo_norm")]<-"Norm. Chla"
blbl$variable[which(blbl$variable=="bbp_norm1")]<-"Norm. Bbp"
blbl$variable[which(blbl$variable=="ration")]<-"N. Chla / N. bbp"
blbl$variable<-factor(blbl$variable, levels=c("Norm. Chla","Norm. Bbp","N. Chla / N. bbp"))

bigplot1<-ggplot(subset(blbl, month %in% c(3,4,5,6)), aes(x=depth_norm, y=value))+
      geom_boxplot(aes(group=depth_norm_layers), outlier.size=0.5)+
  theme_bw()+coord_flip(xlim = c(2,0), ylim=c(0,2))+
  facet_grid(monthn~variable, scale="free")+
  geom_vline(xintercept  = 1)+xlab('Depth - norm.')+ylab('')

bigplot2<-ggplot(subset(blbl, month %in% c(7,8,9,10)), aes(x=depth_norm, y=value))+
      geom_boxplot(aes(group=depth_norm_layers), outlier.size=0.5)+
  theme_bw()+coord_flip(xlim = c(2,0), ylim=c(0,2))+
  facet_grid(monthn~variable, scale="free")+
  geom_vline(xintercept  = 1)+xlab('Depth - norm.')+ylab('')
```

@@

ggplot(tmp_sig, aes(x=month, fill=variable, color=variable)) +
    geom_smooth(aes(y=value),stat = 'summary',
#              fun.data = median_hilow,
              fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), alpha=0.5)+
  geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2, notch = TRUE)+
        scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
  theme_bw()+
  theme(panel.grid.minor.x = element_line(colour = "grey"),
        panel.grid.major.x = element_line(colour = "white")) +
  ylab('[kg/m3]')+coord_cartesian(ylim=c(16,10))+scale_y_reverse()+
      scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal,
                       name="Density Horizons", 
                       labels=list(TeX('$\\sigma_{MLD}$'),
                                TeX('$\\sigma_{DCM}$'),
                                TeX('$\\sigma_{low}$'),
                                TeX('$\\sigma_{50,bottom}$'),
                                TeX('$\\sigma_{50,top}$'),
                                TeX('$\\sigma_{\\mathrm{PAR}\\,1%}$')))

@@

#relative light
#tmp_iso1 <- subset(tmp_fitco, select = c('month','ISOMLD','ISOMAX','ISOlow','ISO90B','ISO90T','ISO_iso1' ,'dcm_or_not','year','shape'))
tmp_iso1 <- subset(tmp_fitco, select = c('month','ISOMLD','ISOMAX','ISOlow','ISO90B','ISO90T','ISO_iso1','dcm_or_not','year'))

tmp_iso1$ISOMAX[tmp_iso1$dcm_or_not==0]<-NA
tmp_iso1$ISOMAX[(tmp_iso1$month<3) | (tmp_iso1$month>10)]<-NA

tmp_iso <- melt(tmp_iso1[,-which(colnames(tmp_iso1)=='dcm_or_not')], id.vars=c('month','year'))

relative_light <- ggplot(tmp_iso, aes(x=month, fill=variable, color=variable)) +
    geom_smooth(aes(y=value),stat = 'summary',
              fun.args = list(conf.int = 0.95), alpha=0.5)+
  geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2, notch = TRUE)+
        scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
theme_bw()+
  theme(panel.grid.minor.x = element_line(colour = "grey"),
        panel.grid.major.x = element_line(colour = "white")) +
  ylab('% of Incoming PAR') +  scale_y_continuous(trans = 'log10') +
  scale_color_manual(aesthetics = c("colour", "fill"),
                         values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E", "#E6AB02"),
                       name="ISO Horizons", 
                       labels=list(TeX('$\\ISO_{MLD}$'),
                                   TeX('$\\ISO_{DCM}$'),
                                   TeX('$\\ISO_{low}$'),
                                   TeX('$\\ISO_{50,bottom}$'),
                                   TeX('$\\ISO_{50,top}$'),
                                   TeX('$\\ISO_{1%}$')))

relative_light

#absolute light
tmp_iso1 <- subset(tmp_fitco, select = c('month','ISOaMLD','ISOaMAX','ISOalow','ISOa90B','ISOa90T','dcm_or_not','year','shape'))

tmp_iso1$ISOaMAX[tmp_iso1$dcm_or_not==0]<-NA
tmp_iso1$ISOaMAX[(tmp_iso1$month<3) | (tmp_iso1$month>10)]<-NA
tmp_iso <- melt(tmp_iso1[,-which(colnames(tmp_iso1)=='dcm_or_not')], id.vars=c('month','year','shape'))

absolute_light <- ggplot(tmp_iso, aes(x=month, fill=variable, color=variable)) +
    geom_smooth(aes(y=value),stat = 'summary',
#              fun.data = median_hilow,
#              fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), alpha=0.5)+
  geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2, notch = TRUE)+
        scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
  theme_bw()+
  theme(panel.grid.minor.x = element_line(colour = "grey"),
        panel.grid.major.x = element_line(colour = "white")) +
  scale_y_continuous(trans = 'log10') +
  ylab(TeX('$\\mu molQuanta\\,m^{-2}\\,s^{-1}]$')) +
  scale_color_manual(aesthetics = c("colour", "fill"),
                     values = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E"),
                       name="ISO Horizons", 
                       labels=list(TeX('$\\ISO_{MLD}$'),
                                   TeX('$\\ISO_{DCM}$'),
                                TeX('$\\ISO_{low}$'),
                                TeX('$\\ISO_{50,bottom}$'),
                                TeX('$\\ISO_{50,top}$')))

absolute_light

# # https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/
# library(RColorBrewer)
# display.brewer.pal(n = 8, name = 'Dark2')
# brewer.pal(n=8, name='Dark2')

#grid.arrange(relative_light + theme(legend.position="none"), absolute_light, nrow = 2)
#grid.arrange(relative_light, absolute_light, nrow = 2)
```
